Silent pessimists - researchers' perceptions of malaria eradication and implications for health spending

### 2017-10-17 paper.Rmd
paper update

In recent years, much of the discourse regarding malaria has shifted back to global eradication [@Roberts2007], with funders, researchers, and public health practitioners rallying to the cause [@Tanner2015]. The Bill and Melinda Gates Foundation has begun actively promoting eradication as feasible "within a generation" [@Gates2014], and the leadership of the World Health Organization (WHO) has supported this goal. Even in areas of high endemicity, advances in immunology, parasitology, modeling and vaccinology, along with rapid economic development, have made eradication appear a more feasible goal, even if not possible in the short term [@Snow2015, @Eckhoff2014]. From both administrative [@Yamey2004] and scientific [@Alonso2011] points of view, eradication has never before received so much attention, nor appeared so within grasp.

@@

Rmisc::multiplot(g1, g3, g5,
                 g2, g4, g6,
                 cols = 2)
```

>>>>

In recent years, much of the discourse regarding malaria has shifted back to global eradication [@Roberts2007], with funders, researchers, and public health practitioners rallying to the cause [@Tanner2015]. The Bill and Melinda Gates Foundation has begun actively promoting eradication as feasible "within a generation" [@Gates2014], and the leadership of the World Health Organization (WHO) has supported this goal, stating the objective of eliminating malaria in 65 new countries from 2015 through 2030 (see below).

```{r}
# Read WHO data (compiled manually from WHO website)
who <- readr::read_csv('who_elimination_data.csv')
who$year <- as.numeric(substr(who$year, 6, 9))
projection <- data.frame(year = seq(2030.2, 2035, length = 20),
                         countries = seq(33, 0, length = 20))
g <- ggplot(data = who %>% mutate(past = ifelse(past, 'Historical',
                                           'WHO objective')) %>%
         filter(!is.na(with_malaria)) ,
       aes(x = year,
           y = with_malaria,
           color = past)) +
  geom_line() +
  geom_point() +
  theme_paper() +
  labs(x = '',
       y = 'Countries',
       title = 'Countries with malaria: Observed and WHO objectives',
       subtitle = 'Data from http://www.who.int/malaria/areas/elimination/overview/en/') +
  scale_color_manual(name = '',
                     values = c('darkorange', 'blue')) +
  ylim(0, 130) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_line(data = projection,
            aes(x = year,
                y = countries,
                color = NA,
                group = NA),
            color = 'black',
            lty = 2) +
  geom_hline(yintercept = 0, col = 'red')
g
# #Make animation for presentation
# make_animation <- function(){
#   g <- 
#     ggplot(data = who %>% mutate(past = ifelse(past, 'Historical',
#                                            'WHO objective')) %>%
#          filter(!is.na(with_malaria)) ,
#        aes(x = year,
#            y = with_malaria,
#            color = past)) +
#   geom_line() +
#   geom_point() +
#   theme_black() +
#   labs(x = '',
#        y = 'Countries',
#        title = 'Countries with malaria: Observed and WHO objectives',
#        subtitle = 'Data from http://www.who.int/malaria/areas/elimination/overview/en/') +
#   scale_color_manual(name = '',
#                      values = c('darkorange', 'blue')) +
#   ylim(0, 130) +
#   theme(axis.text.x = element_text(angle = 90)) +
#   geom_hline(yintercept = 0, col = 'red') +
#     xlim(1970, 2060)
#   add_zero <- function (x, n) {
#     x <- as.character(x)
#     adders <- n - nchar(x)
#     adders <- ifelse(adders < 0, 0, adders)
#     for (i in 1:length(x)) {
#         if (!is.na(x[i])) {
#             x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""), 
#                 x[i], collapse = "")
#         }
#     }
#     return(x)
# }
#   counter <- 0
#   a <- (seq(1,5, length = 15))^2.5
#   a <- round(a)
#   a <- c(a, rev(a) +2)
#   for (i in a){
#     message(counter)
#     counter <- counter +1
#     projection <- data.frame(year = seq(2017, 2035+i, length = 20),
#                          countries = seq(100, 0, length = 20))
#     g <- g + geom_line(data = projection,
#             aes(x = year,
#                 y = countries,
#                 color = NA,
#                 group = NA),
#             color = 'grey',
#             alpha = 0.1) 
#     x <- g +
#   geom_line(data = projection,
#             aes(x = year,
#                 y = countries,
#                 color = NA,
#                 group = NA),
#             color = 'white',
#             lty = 2) 
#     print(x)
#     ggsave(filename = paste0('animation/',
#                              add_zero(counter, n = 3),
#                              '.png'))
#   }
# }
# make_animation()
```

Even in areas of high endemicity, advances in immunology, parasitology, modeling and vaccinology, along with rapid economic development, have made eradication appear a more feasible goal, even if not possible in the short term [@Snow2015, @Eckhoff2014]. From both administrative [@Yamey2004] and scientific [@Alonso2011] points of view, eradication has never before received so much attention, nor appeared so within grasp.

@@

Rmisc::multiplot(g1, g3, g5,
                 g2, g4, g6,
                 cols = 2)

# z <- ggplot(data = who %>% mutate(past = ifelse(past, 'Historical',
#                                            'WHO objective')) %>%
#          filter(!is.na(with_malaria)) ,
#        aes(x = year,
#            y = with_malaria,
#            color = past)) +
#   geom_line() +
#   geom_point() +
#   theme_paper() +
#   labs(x = '',
#        y = 'Countries',
#        title = 'Countries with malaria: Observed and WHO objectives',
#        subtitle = 'Data from http://www.who.int/malaria/areas/elimination/overview/en/') +
#   scale_color_manual(name = '',
#                      values = c('darkorange', 'blue')) +
#   ylim(0, 130) +
#   theme(axis.text.x = element_text(angle = 90)) +
#   geom_line(data = projection,
#             aes(x = year,
#                 y = countries,
#                 color = NA,
#                 group = NA),
#             color = 'white',
#             lty = 2) +
#   geom_hline(yintercept = 0, col = 'red')
# 
# # xa_plot <- xa %>% mutate(year = year + 2016) %>% filter(year == 2066)
# xa_plot <- data.frame(year = c(2015, 2066),
#                       val = c(100, 0))
#   
# z + geom_line(data = xa_plot,
#               aes(x = year,
#                   y = val),
#               color = 'white',
#               alpha = 0.6,
#               group = 1) +
#   theme_black()
```
